The impact of forest management on the temperature sensitivity of organic matter

### 2024-08-01 Manuscript_draft/Draft.Rmd
Updated draft

---
title: "Draft"
author: "Lorenzo Menichetti"
date: "2024-07-04"
output: pdf_document
  #   citation_package: natbib
bibliography: references.bib
header-includes:
   - \usepackage[utf8]{inputenc}
   - \usepackage{textcomp}
   - \usepackage{amsmath}
   - \DeclareUnicodeCharacter{03C9}{$\omega$}  # Declare the Unicode character for omega
---

@@

## Machine learning variance decomposition
On a more limited subset (only Spain), we tested a more extensive random forest model in order to test the predictive power of other variables on the respiration. The model had structure as:  
$$
\begin{aligned}
Rh(t) \sim f(trees, DBH, BD, por, pH, P, N_{tot}, C{tot}, biom_{fungi}, biom_{bact}, \\
biom_{act}, biom_{gram+}, biom_{gram-}, biom_{tot}, T(t), theta_v(t), treatment)
\end{aligned}
$$
We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), bulk density ($BD$), porosity ($por$), pH, phosophorous ($P$), total nitrogen ($N_{tot}$), total carbon ($C_{tot}$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), as well as variables mentioned above.
The aim of this model was to test the predictive potential of any of these additional variables, to explore eventual correlations and possible related processes. The choice of a random forest model over a more conventional linear model is motivated by the capacity of the former to consider more complex relationships and interactions, on top of the convenience of being able to assess quantitatively the predictive potential of each variable by estimating relative variable importance.

@@

## Machine learning variance decomposition

![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition.png){width=60%}

@@

@@

>>>>

---
title: "The impact of forest management on the temperature sensitivity of organic matter"
author: "Leticia Perez-Izquiero, Lorenzo Menichetti, Jorge Curiel-Juste, etc"
date: "2024-07-04"
output:
  bookdown::pdf_document2: default
  bookdown::html_document2: default
  #output: pdf_document
  #   citation_package: natbib
bibliography: references.bib
header-includes:
   - \usepackage[utf8]{inputenc}
   - \usepackage{textcomp}
   - \usepackage{amsmath}
   - \DeclareUnicodeCharacter{03C9}{$\omega$}  # Declare the Unicode character for omega
---

@@

## Machine learning variance decomposition
On a more limited subset (only Spain), we tested a more extensive random forest model in order to test the predictive power of other variables on the respiration. The model had structure as:  
$$
\begin{aligned}
Rh(t) \sim f(trees, DBH, BD, por, pH, P, N_{tot}, C{tot}, biom_{fungi}, biom_{bact}, \\
biom_{act}, biom_{gram+}, biom_{gram-}, biom_{tot}, T(t), theta_v(t), treatment)
\end{aligned}
$$
We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), bulk density ($BD$), porosity ($por$), pH, phosophorous ($P$), total nitrogen ($N_{tot}$), total carbon ($C_{tot}$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), as well as variables mentioned above.  
This model was tested only on a subset of the dataset (only the sites in Spain), for which all the additional variables were available.
The aim of this model was to test the predictive potential of any of these additional variables, in order to to explore eventual correlations and possible related processes. The choice of a random forest model over a more conventional linear model is motivated by the capacity of the former to consider more complex relationships and interactions, on top of the convenience of being able to assess quantitatively the predictive potential of each variable by estimating relative variable importance.

@@

## Machine learning variance decomposition
The fitness of the detailed random forest model (fitted only on data from Spain) was the highest of all the models we tested, unsurprisingly given the huge amount of predictors compared with all other models.   
Soil temperature and moisture resulted the most important variables in predicting respiration from each site by far, followed by edaphic charachrteristics of the site (density and porosity), then by treatment, and then again by edaphic site characteristics (namely pH and total C). Only afterwards the first variable related to aboveground vegetation. Variables related to biomass, total and specific to microorganisms classes, come only later and did not seem particularly important.  
![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition.png){width=50%}

@@

## Role of temperature and moisture in conditioning decomposition
Temperature and moisture seem the most important components in explaining the variance of the soil respiration. The seasonality scaling, although apparently differentiating slightly between sites, presents always very low amplitude posteriors, weighting on average only a 5% of the total measured respiration. It was more important in France (between 4.5% and 12.2%) and much less in Spain and Romania (between 2.0% and 3.2%). This seasonality might represent many different processes, but the most likely are seasonal processes connected with the above-ground vegetation and their inputs to the soil, which in turn would control microbial communities.

## Activation energy of different organic material

## Role of edaphic, ecological and microbiological variables in conditioning decomposition
Any other variable than soil temperature and moisture did not seem to play a pivotal role in explaining the variance of the soil respiration.

## Consequences of different temperature sensitivity of organic material in reacting to warming

@@

![Ratio (expressed as %) between the mean amplitude parameter and the mean measured respiration by treatment](../Figures/Amplitude_ratio.png)
